Finiteness in Cubical Type Theory

### 2020-07-05 sections/search.tex
bit more

Again, there's not any great revelation in ease of use here, but more complex
examples really benefit.
Especially when we build the full set of instances: any expression built out of
products and sums will automatically have an instance.
This will allow us, for instance, to perform proof search over tuples, which
gives us some degree of automation for proof search in tuples.
\begin{agdalisting}
  \ExecuteMetaData[agda/Snippets/Bool.tex]{and-comm}
\end{agdalisting}
\subsection{Generic Currying and Uncurrying}
While we have arguably removed the bulk of the boilerplate from the automated
proofs, there is still the case of the ugly noise of currying and uncurrying.
In this section, we take inspiration from
\cite{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
\(n\)-ary functions and properties.
What we provide here differs from that work in the following ways:
\begin{itemize}
  \item Our generic representation can handle dependent \(\sum\) and \(\prod\)
    types (rather than their non-dependent counterparts, \(\times\) and
    \(\rightarrow\)).
    This extension was necessary for our use case: it is mentioned in the paper
    as the obvious next step. 
  \item We implement the curry-uncurry combinators as (verified) isomorphisms.
    Since we are in a cubical setting, this gives us equivalences between the
    types, a feature not available in standard Agda.
  \item We deal with implicit and instance arguments generically.
\end{itemize}
A full explanation of our implementation is beyond the scope of this work, but
we will mention the key parts here.
First, we define a function arrow generic over the application method:
% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-arr}
Then, we prove that it is isomorphic to the normal function arrow:
% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-iso}
This step will allow us to write the curry-uncurry proofs once, and then extend
them to the three different argument forms without difficulty.

>>>>

Again, there's not any great revelation in ease of use here, but more complex
examples really benefit.
Especially when we build the full set of instances: any expression built out of
products and sums will automatically have an instance.
This will allow us, for instance, to perform proof search over tuples, which
gives us some degree of automation for proof search in tuples.
\begin{agdalisting}
  \ExecuteMetaData[agda/Snippets/Bool.tex]{and-comm}
\end{agdalisting}
\subsection{Generic Currying and Uncurrying}
While we have arguably removed the bulk of the boilerplate from the automated
proofs, there is still the case of the ugly noise of currying and uncurrying.
In this section, we take inspiration from
\citet{allaisGenericLevelPolymorphic2019} to develop a small interface to generic
\(n\)-ary functions and properties.
We will describe it briefly here.

The basic idea of currying and uncurrying generically is to allow ourselves to
work with a generic and flexible representation of function arguments which can
be manipulated more easily than a simple function itself.
Our first task, then, is to define that representation of function arguments.
As in \citet{allaisGenericLevelPolymorphic2019}, our representation is a tuple
which is in some sense a ``second order'' indexed type.
By second order here we mean that it is an indexed type indexed by another
indexed type.
The reason for this complexity is that our solution is to be fully
level-polymorphic.
To start, we define a type representing a vector of universe levels:

\begin{minipage}{\linewidth}
  \begin{multicols}{2}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Data/Product/NAry.tex]{levels}
    \end{agdalisting} \columnbreak
    \begin{agdalisting}
      \ExecuteMetaData[agda/Data/Product/NAry.tex]{max-level}
    \end{agdalisting}
  \end{multicols}
\end{minipage}

This will be used to assign our tuple the correct universe level generically.
Next, we define the list of types (this type is indexed by the list of universe
levels of each type):
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/Product/NAry.tex]{types}
\end{agdalisting}
And finally, the tuple, indexed by its list of types:

\begin{minipage}{\linewidth}
  \begin{multicols}{2}
    \begin{agdalisting}
      \ExecuteMetaData[agda/Data/Product/NAry.tex]{tuple-plus}
    \end{agdalisting} \columnbreak
    \begin{agdalisting}
      \ExecuteMetaData[agda/Data/Product/NAry.tex]{tuple}
    \end{agdalisting}
  \end{multicols}
\end{minipage}
The reason for two separate functions here is to avoid the
\(\AgdaDatatype{\(\top\)}\)-terminated tuples we would need if we just had one.
This means that, for instance, to represent a tuple of a \AgdaDatatype{Bool} and
\agdambb{N} we can write
\((\AgdaInductiveConstructor{true}\;\AgdaInductiveConstructor{,}\;\AgdaNumber{2})\)
instead of 
\((\AgdaInductiveConstructor{true}\;\AgdaInductiveConstructor{,}\;\AgdaNumber{2}\;\AgdaInductiveConstructor{,}\;\AgdaInductiveConstructor{tt})\).

Next we turn to how we will represent functions.
In Agda there are three ways to pass function arguments: explicitly,
implicitly, and as an instance.
We will represent these three different versions with a data type:
\begin{agdalisting}
  \ExecuteMetaData[agda/Data/Product/NAry.tex]{arg-form}
\end{agdalisting}

What we provide here differs from that work in the following ways:
\begin{itemize}
  \item Our generic representation can handle dependent \(\sum\) and \(\prod\)
    types (rather than their non-dependent counterparts, \(\times\) and
    \(\rightarrow\)).
    This extension was necessary for our use case: it is mentioned in the paper
    as the obvious next step. 
  \item We implement the curry-uncurry combinators as (verified) isomorphisms.
    Since we are in a cubical setting, this gives us equivalences between the
    types, a feature not available in standard Agda.
  \item We deal with implicit and instance arguments generically.
\end{itemize}
A full explanation of our implementation is beyond the scope of this work, but
we will mention the key parts here.
First, we define a function arrow generic over the application method:
% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-arr}
Then, we prove that it is isomorphic to the normal function arrow:
% \ExecuteMetaData[agda/Data/Product/N-ary.tex]{fun-iso}
This step will allow us to write the curry-uncurry proofs once, and then extend
them to the three different argument forms without difficulty.
